`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks

### 2020-02-08 analysis/paper/sim_survey_paper.Rmd
Revise paper up to the Core functions section (add details on the age sampling and analysis changes)

This simulation framework has similarities to those presented by Schnute and Haigh [-@schnute2003] and Puerta et al. [-@puerta2019], however, efforts were focused on developing a series of general and accessible functions to simplify the process of testing multiple sampling scenarios and analytical pathways. The steps taken to simulate surveys of spatial, age-structure populations are outlined below. First we outline the equations underlying the package in the **[Model structure]** section and then move into a vignette of the package under the **[Using **`SimSurvey`**]** section. While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study. Output from default function calls are therefore relevant to the case study and these results are described and discussed in **[S1 Appendix][S1 Appendix: Case study]**.

@@

In this section, we describe the framework currently implemented in the **`SimSurvey`** package. With this framework, we tried to strike a balance between realism, simplicity, generality and computational feasibility. The framework follows four general steps: 1) simulate a spatially-aggregated age-structured population; 2) distribute the population throughout a spatial grid, imposing correlation across space, time and age; 3) sample the population using random sampling; and, 4) obtain population estimates using a design-based analysis. Though there is a degree of flexibility in each of these steps, users can circumvent specific components by applying user defined equations, inputs and/or analyses. Details on how to use the package and, if desired, circumvent some aspects of its structure are outlined in the **[Using **`SimSurvey`**]** section.

@@

Rather than developing a full spatially-explicit model, population and spatial dynamics are modeled as independent processes for simplicity. The complexities of spatial population dynamics - such as larval dispersal, spatial differences in growth and population connectivity - are not explicitly accounted for and, as such, the model is a necessary simplification of reality. Despite this limitation, the approach taken facilitates the simulation of spatial, age-structured populations with sufficient complexity for testing the efficacy of various survey designs. The simplicity also limits the number of unknown parameters that need to be specified to simulate a population. Parameter estimates from  spatially-aggregated age-structured models, which are commonly used in stock assessments, can therefore be used to simulate a population using the cohort model and the resultant abundance at age values can be distributed across a spatial grid. Here, a grid of $s$ cells is generated where each cell has an area of $A$ and depth $d$; depth is defined using a sigmoid curve, applied across one spatial axis, with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$. We use depth as our main stratification variable, but note that any other appropriate stratification variables could be used. The grid can be divided into two hierarchical levels, management divisions and habitat-based survey strata.  For demonstration purposes, we envision these levels as part of a stratified-random survey within international fishery divisions, i.e., $H_{\mathrm{strat}}$ depth-based strata within $H_{\mathrm{div}}$ divisions (e.g. NAFO or ICES divisions, or any other geographically bounded area). The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space. This noise term is scaled to sum to 1 to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model:

@@

Once catches are simulated, lengths of the fish sampled by each set are simulated using the von Bertalanffy growth equation found above in the **[Simulate abundance]** section. Sub-sampling is then conducted whereby a subset of fish are sampled for length measurements and a subset of this subset are sampled for age determination. Specifically, a maximum number of lengths are measured per set, $M_{\mathrm{lengths}}$, and a maximum number of ages, $M_{\mathrm{ages}}$, are sampled per length group, $l_{\mathrm{group}}$, per division. Such sub-sampling is common in fisheries-independent surveys as it is costly, impractical and unnecessary to sample every fish captured. Age determination is especially time-consuming, which is why otoliths for age-determination tend to be sub-sampled by length-bin to obtain a representative age sample across a wider range of lengths than would be obtained via random sampling.

@@

While there are many model-based options for obtaining an abundance index from survey data (e.g. [@Thorson2015]), design-based approaches, such as stratified analyses, are often used. Here we apply formula presented in Smith and Somerton [-@Smith1981; equations are replicated in **[S3 Appendix][S3 Appendix: Stratified analysis equations]**] to calculate stratified estimates of total abundance ($\hat{I}_{y,k}$), abundance at length ($\hat{I}_{l,y,k}$) and abundance at age ($\hat{I}_{a,y,k}$). Note that estimates of total abundance are based on aggregate catch while abundance at length requires length frequencies to be scaled up using set-specific ratios of measured to caught fish, and these length frequencies are converted to age by applying a division-level age-length-key. We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance at age estimates from each survey:

$$RMSE = \sqrt{\frac{ \sum_{a = 1}^{N_a} \sum_{y = 1}^{N_y} \sum_{k = 1}^{N_k} (\hat{I}_{a,y,k} - I_{a,y})^2}{N_a N_y N_k}}$$

Where $N_a$, $N_y$, and $N_k$ are the number of ages, years and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e. catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$). RMSE was also calculated for abundance at length estimates, where the above formula is indexed by length groups $l$, and total abundance, which lacks a group index of $a$ or $l$.

>>>>

This simulation framework has similarities to those presented by Schnute and Haigh [-@schnute2003] and Puerta et al. [-@puerta2019], however, efforts were focused on developing a series of general and accessible functions to simplify the process of testing multiple sampling scenarios and analytical pathways. The steps taken to simulate surveys of spatial, age-structure populations are outlined below. Under the **[Methods]** section, we first outline the equations underlying the package (**[Model structure]** section) and then move into a vignette of the package (**[Core functions]** section). While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study. Output from default function calls are therefore relevant to the case study and some of these results are highlighted in the **[Results]** and **[Discussion]** section; a more detailed description and discussion of these results can be found in **[S1 Appendix][S1 Appendix: Case study]**.

@@

In this section, we describe the framework currently implemented in the **`SimSurvey`** package. With this framework, we tried to strike a balance between realism, simplicity, generality and computational feasibility. The framework follows four general steps: 1) simulate a spatially-aggregated age-structured population; 2) distribute the population throughout a spatial grid, imposing correlation across space, time and age; 3) sample the population using random sampling; and, 4) obtain population estimates using a design-based analysis. Though there is a degree of flexibility in each of these steps, users can circumvent specific components by applying user defined equations, inputs and/or analyses. Details on how to use the package and, if desired, circumvent some aspects of its structure are outlined in the **[Core functions]** section.

@@

Rather than developing a full spatially-explicit model, population and spatial dynamics are modeled as independent processes for simplicity. The complexities of spatial population dynamics - such as larval dispersal, spatial differences in growth and population connectivity - are not explicitly accounted for and, as such, the model is a necessary simplification of reality. Despite this limitation, the approach taken facilitates the simulation of spatial, age-structured populations with sufficient complexity for testing the efficacy of various survey designs. The simplicity also limits the number of unknown parameters that need to be specified to simulate a population. Parameter estimates from  spatially-aggregated age-structured models, which are commonly used in stock assessments, can therefore be used to simulate a population using the cohort model and the resultant abundance at age values can be distributed across a spatial grid. Here, a grid of $s$ cells is generated where each cell has an area of $S$ and depth $d$; depth is defined using a sigmoid curve, applied across one spatial axis, with a depth range of $[d_{\mathrm{min}}, d_{\mathrm{max}}]$, shelf depth of $d_{\mathrm{shelf}}$ and a shelf width of $w_{\mathrm{shelf}}$. We use depth as our main stratification variable, but note that any other appropriate stratification variables could be used. The grid can be divided into two hierarchical levels, management divisions and habitat-based survey strata.  For demonstration purposes, we envision these levels as part of a stratified-random survey within international fishery divisions, i.e., $H_{\mathrm{strat}}$ depth-based strata within $H_{\mathrm{div}}$ divisions (e.g. NAFO or ICES divisions, or any other geographically bounded area). The simulated population is distributed through the grid by simulating spatial-temporal noise controlled by a parabolic relationship with depth and covariance between ages, years and space. This noise term is scaled to sum to 1 to ensure that the total population of each age for each year through the grid equals the number simulated by the cohort model:

@@

Once catches are simulated, lengths of the fish sampled by each set are simulated using the von Bertalanffy growth equation found above in the **[Simulate abundance]** section. Sub-sampling is then conducted whereby a subset of fish are sampled for length measurements and a subset of this subset are sampled for age determination. Specifically, a maximum number of lengths are measured per set, $M_{\mathrm{lengths}}$, and a maximum number of ages, $M_{\mathrm{ages}}$, are sampled per length group, $l_{\mathrm{group}}$, per division, strata or set. Such sub-sampling is common in fisheries-independent surveys as it is costly, impractical and unnecessary to sample every fish captured. Age determination is especially time-consuming, which is why otoliths for age-determination tend to be sub-sampled by length-bin to obtain a representative age sample across a wider range of lengths than would be obtained via random sampling.

@@

While there are many model-based options for obtaining an abundance index from survey data (e.g. [@Thorson2015]), design-based approaches, such as stratified analyses, are often used. Here we apply formula presented in Smith and Somerton [-@Smith1981] (equations are replicated in **[S3 Appendix][S3 Appendix: Stratified analysis equations]**) to calculate year $y$ and simulation $j$ specific stratified estimates of total abundance ($\hat{I}_{y,j}$), abundance at length ($\hat{I}_{l,y,j}$) and abundance at age ($\hat{I}_{a,y,j}$). Note that estimates of total abundance are based on total numbers caught at each set, $n_i$, while abundance at length requires the sub-sampled length frequencies at each set, $m_{l, i}$, to be scaled up using set-specific ratios of the number of fish measured, $m_{i}$, to numbers caught, $n_i$:

$$n_{l, i} = m_{l, i} \times \frac{n_i}{m_i}$$

Likewise, age frequencies need to be calculated to obtain stratified estimates of abundance at age. This is done by constructing an age-length key, which is the proportion of fish in each length bin that fall into specific ages classes, and applying this key to the bumped up length frequencies, $n_{l,i}$, to approximate age frequencies, $n_{a,i}$, that is:

$$n_{a,i} = \sum_{l = 1}^{L} n_{l,i} \times \frac{m^{\prime}_{a,l,g}}{m^{\prime}_{l,g}}$$

Here, the prime symbol $^{\prime}$ indicates that these values are tertiary sampling units as they represent fish sub-sampled for age determination from those sub-sampled for length measurements. Within this sub-sample, $m^{\prime}_{l,g}$ represents the number of fish within length bin $l$ and from area $g$, and $m^{\prime}_{a,l,g}$ represents the number of fish within age-class $a$, length bin $l$ and area $g$. The area $g$ can be division, strata or set specific. Though age-length keys are typically constructed and applied over large spatial scales (e.g. division), calculating proportions at finer scales may better account for the hierarchical sample designs [@Aanes2015].

We used root-mean-squared error (RMSE) as a measure of the precision and bias of the abundance at age estimates from each survey:

$$RMSE = \sqrt{\frac{ \sum_{a = 1}^{A} \sum_{y = 1}^{Y} \sum_{j = 1}^{J} (\hat{I}_{a,y,j} - I_{a,y})^2}{A \times Y \times J}}$$

Where $A$, $Y$, and $J$ are the number of ages, years and simulations, respectively, and $I_{a,y}$ is the true abundance available to the survey (i.e. catchability corrected abundance; $I_{a,y} = q_a N_{a,y}$). RMSE was also calculated for abundance at length estimates, where the above formula is indexed by length groups $l$, and total abundance, which lacks a group index of $a$ or $l$.
